COLLISION RATES OF PLANETESIMALS NEAR MEAN-MOTION RESONANCES

### 2019-12-10 main.tex
Update timestepping description

Under the effect of MMR, a body will undergo oscillations in semi-major axis and eccentricity. So long as the gravitational force on 
the body is dominated by the central star and the perturbing body, the amplitude of the oscillations in semi-major axis can be 
described using the pendulum approximation (see equation 6 in \citet{2019MNRAS.489.2159W}). This amplitude can be thought of 
as the 'width' over which the perturbations of the resonance are effective. Variations in eccentricity follow the semi-major axis 
variations according to the Tisserand relation

@@

\noindent with $\alpha$ = 3/2 and $\Sigma_{0}$ = 10 g cm$^{-2}$. Planetesimals are given a bulk density of 2 g cm$^{-3}$ and a 
diameter of 300 km. This corresponds to a disk containing slightly less than 500,000 planetesimals. Both simulations are evolved for 
10,000 years, which is about 800 Jupiter orbits.

In the CJ simulation, there is no secular forcing from Jupiter and so the mean longitudes and mean anomalies are randomly 
assigned values between 0 and 2$\pi$. The distribution of eccentricities and inclinations of the planetesimals are chosen so that the 
effects of viscous stirring and gas drag from the solar nebula are in balance. This occurs when eccentricities are described by a 
Rayleigh distribution with a scale parameter given by equation 12 in \citet{2002ApJ...581..666K}. The inclination distribution is 
described by this same function, with a scale parameter that is half as large. This is typical for a sufficiently dynamically heated disk 
of particles in a Kepler potential \citep{1993MNRAS.263..875I}.

For simulation EJ, the effects of secular forcing by Jupiter are applied to the initial conditions. This is done by first assigning Kepler 
orbital elements to the planetesimals in the same way as simulation CJ and then adding a forced component to the eccentricity 
vectors $z_{i} = \left( h_{i}, k_{i} \right)$, described by equation \ref{eq:eforced}. For simplicity, a coordinate system is chosen so that 
$\omega_{g}$ = 0. This allows us to simply add a scalar value to each $h_{i}$, which depends only on the semi-major axis of a 
planetesimal and Jupiter, along with the eccentricity of Jupiter.

@@

For the purposes of the integrator, there are two relevant timescales in this system. The first is the orbital dynamical time $\sqrt{a^3/
G M_{\odot}}$. All particles are placed on a fixed time step of $\Delta T$ = 0.01 yr. This is roughly 3\% of an orbital dynamical time at 
the inner edge of the planetesimal disk, which is a reasonable time step size for a leapfrog integrator (reference?). The time step for 
collisions is determined by the planetesimal dynamical timescale ($\sim 1/\sqrt{G \rho}$), which is about 45 minutes. To resolve both 
of these timescales simultaneously, we use a two-tiered time stepping scheme. At the beginning of each time step, all bodies are 
placed on the orbital time step. A first pass of collision detection is then run in which the radii of all bodies are inflated by a factor of 
2.5. Any bodies with imminent collisions predicted using the inflated radii are placed on a time step that is a factor of 16 smaller than 
the orbital time step. The purpose of two-tiered scheme is to properly resolve the gravitational interactions between any bodies that 
will undergo a close encounter. This prevents the coarser base time step from reducing the effectiveness of gravitational focusing, 
while minimizing the additional computational expense.

>>>>

Under the effect of MMR, a body will undergo oscillations in semi-major axis and eccentricity. So long as the gravitational force on 
the body is dominated by the central star and the perturbing body, the amplitude of the oscillations in semi-major axis can be 
described using the pendulum approximation (see equation 8.76 in \citet{2000ssd..book.....M}). This amplitude can be thought of 
as the 'width' over which the perturbations of the resonance are effective. Variations in eccentricity follow the semi-major axis 
variations according to the Tisserand relation

@@

\noindent with $\alpha$ = 3/2 and $\Sigma_{0}$ = 10 g cm$^{-2}$. Planetesimals are given a bulk density of 2 g cm$^{-3}$ and a 
diameter of 300 km. This corresponds to a disk containing roughly 500,000 planetesimals. Both simulations are evolved for 
5,000 years, which is about 400 Jupiter orbits.

In the CJ simulation, there is no secular forcing from Jupiter and so the mean longitudes and mean anomalies are randomly 
assigned values between 0 and 2$\pi$. The distribution of eccentricities and inclinations of the planetesimals are chosen so that the 
effects of viscous stirring and gas drag from the solar nebula are in balance. This occurs when eccentricities are described by a 
Rayleigh distribution with a scale parameter given by equation 12 in \citet{2002ApJ...581..666K}. The inclination distribution is 
described by this same function, with a scale parameter that is half as large. This is typical for a dynamically heated disk 
of particles in a Kepler potential \citep{1993MNRAS.263..875I}.

For simulation EJ, the effects of secular forcing by Jupiter are built in to the initial conditions. This is done by first assigning Kepler 
orbital elements to the planetesimals in the same way as simulation CJ and then adding a forced component to the eccentricity 
vectors $z_{i} = \left( h_{i}, k_{i} \right)$, described by equation \ref{eq:eforced}. For simplicity, a coordinate system is chosen so that 
$\omega_{g}$ = 0. This allows us to simply add a scalar value to each $h_{i}$, which depends only on the semi-major axis of a 
planetesimal and Jupiter, along with the eccentricity of Jupiter.

@@

For the purposes of the integrator, there are two relevant timescales in this system. The first is the orbital dynamical time $\sqrt{a^3/
G M_{\odot}}$. All particles are placed on a fixed time step of $\Delta T$ = 0.01 yr. This is roughly 3\% of an orbital dynamical time at 
the inner edge of the planetesimal disk. From numerical tests with {\sc ChaNGa}, we have found that this time step size preserves 
the symplectic nature of the integration without being too computationally expensive. Although using $\Delta T$ = 0.01 yr keeps the 
integration symplectic, we find that a small amount of artificial precession is produced, which is particularly noticable in the EJ 
simulation. We reduce the base time step size by an additional factor of 4 to the keep longitude of perihelia of planetesimals near the 
inner edge of the disk from drifting too far over the course of our integration.

An additional timescale is set by the dynamical time of the planetesimals ($\sim 1/\sqrt{G \rho}$), which is about 45 minutes. To 
resolve the base time step and the dynamical timescale of planetesimals simultaneously, we use a two-tiered time stepping scheme. 
At the beginning of each time step, all bodies are placed on the orbital time step. A first pass of collision detection is then run in 
which the radii of all bodies are inflated by a factor of 2.5. Any bodies with imminent collisions predicted using the inflated radii are 
placed on a time step that is a factor of 16 smaller than the orbital time step. The purpose of two-tiered scheme is to properly resolve 
the gravitational interactions between any bodies that will undergo a close encounter. This prevents the coarser base time step from 
reducing the effectiveness of gravitational focusing, while minimizing the additional computational expense.
